---
output:
  pdf_document: default
  html_document: default
---
---
TITLE: "COMPSCIX 415.2 Homework 3"
Author: "Ganesh Saravanan"
Date: "6/25/2018"
Output: pdf_document

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

My Github repository for my assignments can be found at this URL:<https://github.com/gsaravanan1/rstudiodemo.git>

```{r load_packages, warning=FALSE, message=FALSE}
library(mdsr)
library(tidyverse)
library(ggplot2)
library(nycflights13)
```



# 3.8.1 Exercises
#1 What is the problem with this plot? How could you improve it?

```{r plot1}
ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) + 
  geom_point()
```

#ANSWER: Add a small amount of random variation

```{r plot2}
ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) + 
  geom_jitter()
```


#2. What parameters to geom_jitter() control the amount of jittering?

#ANSWER: width and height.

#3. Compare and contrast geom_jitter() with geom_count().

#ANSWER: Jittering is adding a small amount of random noise to data. It is often used to spread out points that would otherwise be overplotted. It is only effective in the non-continuous data case where overplotted points typically are surrounded by whitespace - jittering the data into the whitespace allows the individual points to be seen. It effectively un-discretizes the discrete data.

#4. What’s the default position adjustment for geom_boxplot()? Create a visualisation of the mpg dataset that demonstrates it.

#ANSWER: The default position adjustment is position_dodge().

```{r}
ggplot(data = mpg, mapping = aes(x = class, y = hwy, color = drv)) + 
  geom_boxplot(position = "dodge")
```



#Section 3.9.1: #2 and #4 only

#2 What does labs() do? Read the documentation.

#ANSWER: labs() adds labels to the graph. You can add a title, subtitle, and a label for the x and y axes, as well as a caption.


#4 What does the plot below tell you about the relationship between city and highway mpg? Why is  coord_fixed() important? What does geom_abline() do?

```{r}
ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) +
  geom_point() + 
  geom_abline() +
  coord_fixed()
```


#ANSWER: Highway MPG is always (mostly) better than city MPG. coord_fixed() forces a specified ratio between the physical representation of data units on the axes. geom_abline() draws a line that, by default, has an intercept of 0 and slope of 1.


#Section 4.4: #1 and #2 only

#1 Why does this code not work?

```{r}
my_variable <- 10
#my_varıable
```

#ANSWER: Typo

#2 Tweak each of the following R commands so that they run correctly:
  
```
library(tidyverse)

ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy))

fliter(mpg, cyl = 8)
filter(diamond, carat > 3)
```

#ANSWER:
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy))
  
  filter(mpg, cyl == 8)
  filter(diamonds, carat > 3)
```


#Section 5.2.4: #1, #3 and #4 only. 
  
#1. Find all flights that

#1.1. Had an arrival delay of two or more hours.
filter(flights, arr_delay>=120)

#1.2. Flew to Houston (IAH or HOU)
filter(flights, dest == 'IAH' | dest == 'HOU')

#1.3. Were operated by United, American, or Delta
filter(flights, carrier == 'UA' | carrier == 'AA' | carrier == 'DL')

#1.4. Departed in summer (July, August, and September)
filter(flights, month >= 7 & month <= 9)

#1.5. Arrived more than two hours late, but didn't leave late
filter(flights, arr_delay > 120, dep_delay <= 0)

#1.6. Were delayed by at least an hour, but made up over 30 minutes in flight
filter(flights, dep_delay >= 60, dep_delay-arr_delay > 30)

#1.7. Departed between midnight and 6am (inclusive)
filter(flights, dep_time <=600 | dep_time == 2400)  

#3 How many flights have a missing dep_time? What other variables are missing? What might these rows represent?

```{r}
summary(flights)
```

#ANSWER :
  8255 flights have a missing dep_time, 8255 have a missing dep_delay, 8713 have a missing arr_time, 9430 have a missing arr_delay, and 9430 have a missing air_time. We can speculate that these are flights that failed to depart or arrive, since a flight that departs normally but is then rerouted will probably have a normally recorded departure but no similar record for it's arrival. However, these could also just be lost data about perfectly normal flights.
  
#4 Why is NA ^ 0 not missing? Why is NA | TRUE not missing? Why is FALSE & NA not missing? Can you figure out the general rule? (NA * 0 is a tricky counterexample!)

#ANSWER: NA ^ 0 evaluates to 1 because anything to the power of 0 is 1, so although we didn't know the original value, we know it's being taken to the zeroth power.

With NA | TRUE, since the | operator returns TRUE if either of the terms are true, the whole expression returns true because the right half returns true. This is easier to see in an expression like NA | 5<10 (since 5 is indeed less than 10).

For the next example, we know that & returns TRUE when both terms are true. So, for example, TRUE & TRUE evaluates to TRUE. In FALSE & NA, one of the terms is false, so the expression evaluates to FALSE. As does something like FALSE & TRUE.

NA * 0 could be argued to be because the NA could represent Inf, and Inf * 0 is NaN (Not a Number), rather than NA. However, I suspect that these results are dictated as much by what answer is natural, quick and sensible in C as by mathematical edge cases.



#Section 5.4.1: #1 and #3 only


#1 Brainstorm as many ways as possible to select dep_time, dep_delay, arr_time, and arr_delay from flights.

#ANSWER: 
```{r}
select(flights, dep_time, dep_delay, arr_time, arr_delay)
select(flights, dep_time,  dep_delay, arr_time, arr_delay)
select(flights, c(dep_time,  dep_delay, arr_time, arr_delay))
flights %>% select(dep_time,  dep_delay, arr_time, arr_delay)
flights %>% select_("dep_time",  "dep_delay", "arr_time", "arr_delay")
flights %>% select_(.dots=c("dep_time",  "dep_delay", "arr_time", "arr_delay"))
```


#3 What does the one_of() function do? Why might it be helpful in conjunction with this vector?
#ANSWER :  one_of() allows for subset-matching
```{r}
vars <- c("year", "month", "day", "dep_delay", "arr_delay")
flights %>% select(one_of(vars))
```

  
