---
title: "COMPSCIX 415.2 Homework 7"
author: "Ganesh Saravanan"
date: "7/18/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Exercise 1.
```{r }
train <- read.csv('~/Downloads/train.csv')
dim(train)
```
There are 1,460 observations and 81 columns.

## Exercise 2.
### 1. Visualize the distribution of SalePrice.
```{r message=FALSE}
library(ggplot2)
ggplot(data = train) + geom_histogram(aes(x = SalePrice))
```

### 2. Visualize the covariation between SalePrice and Neighborhood.
```{r}
ggplot(data = train, aes(x = Neighborhood, y = SalePrice)) +
  geom_boxplot(aes(fill= Neighborhood)) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```
### 3. Visualize the covariation between SalePrice and OverallQual.
```{r message=FALSE}
train$OverallQual <- as.factor(train$OverallQual)
ggplot(data = train, aes(x = OverallQual, y = SalePrice)) +
  geom_boxplot(aes(fill= OverallQual)) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

## Exercise 3.
```{r}
library(broom)
fit <- lm(SalePrice ~1 , data=train)
tidy(fit)

```
The coefficient is 180921.2
```{r}
mean(train$SalePrice)
```
The coefficient is the same as the average value of SalePrice.
```{r}
glance(fit)
```
R-square is 0.

## Exercise 4
```{r}
# Now fit a linear regression model using GrLivArea, OverallQual, and Neighborhood as the features.
options(tibble.print_max = Inf)
fit2 <- lm(data = train, SalePrice ~ GrLivArea + OverallQual+ Neighborhood)
tidy(fit2)
```
- How would you interpret the coefficients on GrLivArea and OverallQual?

  GrLivArea and OverallQual are positive related to SalePrice. For 1 unit increases in GrLivArea,
  the SalePrice will increase $51.37. For OverallQual, when the rating increases, the SalePrice increases as wel.
  
- How would you interpret the coefficient on NeighborhoodBrkSide?

  NeighborhoodBrkSide is negative related to SalePrice. If the house is at NeighborhoodBrkSide,
  the SalePrice will be lower by 23951. 
  
- Are the features significant?

  GrLivArea is statistical significant. OverallQual rating from 6-10 are significant while the others are not. Some neighborhoods are significant to the SalePrice while some are not.

- Are the features practically significant?

  Yes they are practically significant.
  
- Is the model a good fit?

```{r}
glance(fit2)
```
The Adjusted R-squared is 0.8115, I think it provides a good fit.
  
## Exercise 6
```{r message = FALSE, warning=FALSE}
library(tibble)
sim1a <- tibble(
x = rep(1:10, each = 3),
y = x * 1.5 + 6 + rt(length(x), df = 2)
)

fit_sim <- lm(data = sim1a, y~x)
tidy(fit_sim)
```
```{r}
glance(fit_sim)
```
The estimated coefficent of x is around 1.7 (the actual coefficent is 1.5), which is not bad. R-square is 0.9, which looks like a good fit.

```{r}
intercept_list <- rep(NA, 6)
coef_list <- rep(NA,6)
rsquare_list <- rep(NA,6)
adj_rsquare_list <- rep(NA,6)
for(i in 1:6){
    sim1a <- tibble(
  x = rep(1:10, each = 3),
  y = x * 1.5 + 6 + rt(length(x), df = 2)
  )

fit_sim <- lm(data = sim1a, y~x)
intercept_list[i] <- tidy(fit_sim)$estimate[1] # the estimate of intercept
coef_list[i] <- tidy(fit_sim)$estimate[2] # the estimate of coefficent for x
rsquare_list[i] <- glance(fit_sim)$r.squared # r square
adj_rsquare_list[i] <- glance(fit_sim)$adj.r.squared # adjusted r square

}

```

Rerun the simulated data 6 times, here is the coefficent for x:
```{r}
coef_list
```

Here is the rsquare:
```{r}
rsquare_list
```
We can see that the coefficent fluctuates around 1.5 and the rsquare varies a lot. Sometimes it provides a good fit, sometimes it does not.

