---
title: "COMPSCIX 415.2 Homework 1"
author: "Ganesh Saravanan"
date: "6/10/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

My Github repository for my assignments can be found at this URL:<https://github.com/gsaravanan1/rstudiodemo.git>

```{r load_packages, warning=FALSE, message=FALSE}
library(mdsr)
library(tidyverse)
```

Load the WorldCities data

```{r WorldCities}
data("WorldCities")
```

QUESTION 1: How many observations and variables are there in this data set? What are
some of the variable names? Type up your answer in your document using complete sentences.

```{r glimpse}
glimpse(WorldCities)
```

ANSWER 1: Observation=23,018 and Variable=10. Some variable names :code,name, latitude, longitude,populate, etc...


QUESTION 2: There is a column called region in the data set. Can you extract this and show
only the unique values?
```{r extract_region}
region_col <- WorldCities$region
unique(region_col)
```


QUESTION 3: Can you extract and show the unique entries from the country column in
WorldCities using one line of code and two %>% operators?

```{r extract_unique_country}
WorldCities %>% select(country) %>% unique()
```


QUESTION 4: Make one more histogram of the population and add a new x-label, a new ylabel,
a new title (use ggtitle('my title')), and change the theme of the plot using theme_bw().

```{r ggplot}
p <-WorldCities %>% ggplot(aes(x = population)) + geom_histogram()
p + labs(x = "Population", y = "Population_count", title = "City Population Histogram") + theme_bw()
```